\begin{compactenum}
\item [???] \textbf{curl} transfers data to/from a server:
\item [\texttt{I}] fetches the headers only,
\item [\texttt{i}] includes HTTP response headers,
\item [\texttt{L}] redoes requests to relocated pages,
\item [\texttt{o}] writes to a file, not stdin,
\item [\texttt{u}] specifies user name and password,
\item [\texttt{x}] uses proxy,
\item [\texttt{X}] specifies custom request method.
\item [???] \textbf{sftp} is similar to ftp but uses SSH.
\item [???] \textbf{wget} is a network downloader:
\item [\texttt{A}, \texttt{R}] specifies lists of file suffixes or patterns (when wildcard characters appear) to accept/reject,
\item [\texttt{b}] goes to background immediately,
\item [\texttt{c}] continues getting a partially-downloaded file,
\item [\texttt{m}] turns on  mirroring options: infinite recursion and time-stamping,
\item [\texttt{np}] disallows ascension to parent directories,
\item [\texttt{U}] identifies as a different agent to the server.
\item [\texttt{w}] waits specified number of seconds between the retrievals (check: \texttt{--random-wait} too).
\item [???] \textbf{rlogin} starts an unencrypted terminal session on a remote host.
\item [???] \textbf{ssh} executes commands remotely:
\item [\texttt{C}] requests compression of all data,
%\item [\texttt{D}] specifies a local ''dynamic'' application-level port forwarding,
\item [\texttt{i}] uses custom key for authentication,
\item [\texttt{p}] selects a port to connect to on the remote host,
\item [\texttt{X}] enables X11 forwarding.
\end{compactenum}

\begin{compactenum}
\item [???] \textbf{ping} tests reachability of a network host:
\item [\texttt{c}] stops after sending ... packets,
\item [\texttt{n}] no DNS name resolution.
\item [???] \textbf{route} shows and manipulates the IP routing table.
\end{compactenum}

\begin{compactenum}
\item [???] \textbf{rsync} copies (remote or local) files fast:
\item [\texttt{a}] in archive mode, equivalent to:
\item [\texttt{g}] preserves group,
\item [\texttt{o}] preserves owner (super-user only)
\item [\texttt{p}] preserves permissions,
\item [\texttt{t}] preserves modification times,
\item [\texttt{l}] copies symlinks as symlinks,
\item [\texttt{b}] make backups, 
\item [\texttt{c}] skip based on checksum, 
\item [\texttt{n}] performs a dry run without changes made, 
\item [\texttt{r}] resursively,  %%%
\item [\texttt{u}] skip newer files on the receiver, 
\item [\texttt{v}] increases verbosity, %%%
\item [\texttt{z}] compresses file data during the transfer, %%%
\item [???] \textbf{scp} copies files securely (with SSH).
\end{compactenum}
